---
title: "logistic.regression"
---


## Dataset Description

The `efige.txt` dataset contains information on a sample of Italian companies from different economic sectors. Each firm is described by several variables, with their definitions provided in the accompanying legend file.

---

## Modeling Objective

In this section, we'll define and estimate a **logistic regression model** for the binary response variable `widened_prod_range`, which indicates whether a company has expanded its product range. We'll also interpret the effect of each explanatory variable on the likelihood of this outcome.

```{r}
data <- read.delim("efige.txt") 
```

We start by estimating a full logistic regression model that includes fifteen explanatory variables considered theoretically relevant for predicting whether a firm has expanded its product range. 
These variables cover different aspects such as innovation, international openness, ownership structure, geographic location, and firm performance. 
This initial specification serves as the starting point for selecting the most statistically relevant predictors in the next steps.

```{r}
model_full <- glm(widened_prod_range ~ prod_inn + proc_inn + RD_inv +
                  direct_export + import_goods_china_india +
                  group + female_ceo + fam_ceo +
                  north + centre + patent +
                  competitors_from_abroad + qual_cert + deloc_fdi,
                  data = data,
                  family = "binomial")
summary(model_full)
```
From the results of the full model, only a few variables are statistically significant at the conventional 5% level. 

Specifically, `prod_inn`, `proc_inn`, `RD_inv`, `import_goods_china_india`, and `fam_ceo` show a significant association with the likelihood that a firm has expanded its product range.


```{r}
model_final <- glm(widened_prod_range ~ prod_inn + proc_inn + RD_inv +
                   import_goods_china_india + fam_ceo,
                   data = data,
                   family = "binomial")
summary(model_final)
```
The estimated coefficients are positive, suggesting that introducing product or process innovations, investing in R&D, importing intermediate goods from China or India, and having a family CEO all increase the likelihood that a firm has expanded its product range.


```{r}
exp(coef(model_final))
```
#### Interpreting Odds Ratios

The analysis of odds ratios confirms that all selected variables are associated with an increased likelihood of a firm expanding its product range. 

For product innovation, the odds ratio is over 3.8, meaning firms introducing product innovations are more than three times as likely to expand their range. 

Process innovation, R&D investment, imports from China or India, and having a family CEO also show positive effects, with odds ratios between about 1.37 and 1.80.

---

#### Visualizing the Effect of Product Innovation

The following plot shows the percentage of firms that expanded their product range based on whether they introduced product innovations (`prod_inn`). There's a clear difference: innovative firms have a much higher probability of expanding, which aligns with the logistic regression results.

```{r barplot-prod_inn, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)

data %>%
  group_by(prod_inn) %>%
  summarise(perc_widened = mean(widened_prod_range) * 100) %>%
  ggplot(aes(x = factor(prod_inn), y = perc_widened, fill = factor(prod_inn))) +
  geom_col(width = 0.6, show.legend = FALSE) +
  labs(
    x = "Innovazione di prodotto (prod_inn)",
    y = "% imprese con gamma ampliata",
    title = "Ampliamento gamma e innovazione di prodotto"
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Sì")) +
  theme_minimal()
```
As further evidence, the following boxplot shows the distribution of predicted probabilities from the model for the two groups (`prod_inn` = 0 and `prod_inn` = 1). 

Firms that introduced product innovations have noticeably higher fitted values, reinforcing the importance of this variable.

```{r fitted-probabilities-prod_inn, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
# data
data$fitted_probs <- fitted(model_final)

# plot
ggplot(data, aes(x = factor(prod_inn), y = fitted_probs, fill = factor(prod_inn))) +
  geom_boxplot(show.legend = FALSE) +
  labs(
    x = "Product Innovation (prod_inn)",
    y = "Predicted Probability of Expanding Range",
    title = "Predicted Probabilities from Logit Model by prod_inn"
  ) +
  scale_x_discrete(labels = c("0" = "No", "1" = "Yes")) +
  theme_minimal()
```

After examining the effect of individual predictors, we now evaluate the overall fit of the model.

Comparing the null deviance (1544.2) to the residual deviance (1355.3) shows a significant reduction, indicating that the model explains a meaningful portion of the observed variability.

The Akaike Information Criterion (AIC) is 1367.3, which is relatively low and suggests a good balance between explanatory power and simplicity.

Finally, we also calculate the Pseudo R-squared as an additional measure of model quality.


```{r}
pseudo_R2 <- 1 - (model_final$deviance / model_final$null.deviance)
pseudo_R2
```

The final model identifies several significant factors linked to expanding a firm's product range: introducing product and process innovations, investing in R&D, importing intermediate goods from China or India, and having a family CEO.

Overall, the model's fit—based on residual deviance, AIC, and Pseudo R-squared—is satisfactory.

These results align with existing evidence highlighting the importance of innovation, international openness, and governance characteristics in shaping firms' growth strategies.

